Solar wind and geomagnetic activity influence on the equatorial plasma mass density at geosynchronous orbit

### 2016-05-18 paper/paper.tex

\twocolumn[
  \begin{@twocolumnfalse}
\maketitle
\hrule
\begin{abstract}
In this work we consider two types of events, determined by decreases in $D_{st}$ below a threshold value or increases in the $\rho_{eq}$ above a threshold value on 1-hour and 1-day time scales using the \cite{Denton2007} dataset.  From the $D_{st}$ events, we find that there is no statistically significant relationship between either solar wind or geomagnetic activity under conditions of strong geomagnetic activity when hourly or daily averages of $\rho_{eq}$ are considered.  Although very large geomagnetic storms have been observed to correspond to large increases in the plasma trough mass density, in general $\rho_{eq}$ following the onset of a geomagnetic storm does not depend on the storm intensity or the solar wind parameters $B_{z}$ and $P_{dyn}$. From the $\rho_{eq}$ events, we find a weak dependence on $B_{z}$ after the onset of an event, with higher average $B_{z}$ six hours after the event onset corresponding to larger $\rho_{eq}$ events.
\vspace{2em}
\end{abstract}
\hrule
 \end{@twocolumnfalse}
  ]

@@

In the lower panel of Figure \ref{DailyAveragedDstEvents} it is shown that the trends seen in the analysis of 1989-1991 also hold for the entire span of measurements from GOES 6, which spans 1983-1991 - a small elevation on the day of the event relative to the day before and day after, with median averages after the day of the event being slightly lower than that before.

To determine if the magnitude of the density medians on a given day were statistically different than another day, we used a two-population t-test.  For the interval of 1989-1991, corresponding to the top panel of \ref{DailyAveragedDstEvents}, the null hypothesis that any of the medians from days -1, 0, and 1 are different from each other is rejected with a maximum p value of. The results are shown in Table~\ref{Table1}.

These tests indicate that there is not a statistically significant change in medians of $\rho_{eq}$ on the day before and day after a threshold crossing in $D_{st}$.

\input{../tables/DeltaBootstraps.txt}

Caption: Results of test on different between means of $\rho_{eq}$ shown in the top panel of \ref{DailyAverage} on days near or on the day of a $D_{st}$ event.  A difference in the means can be rejected with a probability of false rejection of $p$.

   Days     p
-1 and 0
-1 and 1
 0 and 1

Caption: Results of test on different between means of $\rho_{eq}$ shown in the bottom panel of \ref{DailyAverage} on days near or on the day of a $D_{st}$ event.

>>>>

\twocolumn[
  \begin{@twocolumnfalse}
\maketitle
\hrule
\begin{abstract}
In this work we consider two types of events, determined by decreases in $D_{st}$ below a threshold value or increases in the $\rho_{eq}$ above a threshold value on 1-hour and 1-day time scales using the \cite{Denton2007} dataset.  From the $D_{st}$ events, we find that there is no statistically significant relationship between either solar wind or geomagnetic activity under conditions of strong geomagnetic activity when hourly or daily averages of $\rho_{eq}$ are considered.  Although very large geomagnetic storms have been observed to correspond to large increases in the plasma trough mass density, in general $\rho_{eq}$ following the onset of a geomagnetic storm does not depend on the storm intensity or the solar wind parameters $B_{z}$ and $P_{dyn}$. From the $\rho_{eq}$ events, we find a weak dependence on $B_{z}$ after the onset of an event, with higher average $B_{z}$ six hours after the event onset corresponding to larger $\rho_{eq}$ events.
\end{abstract}
\vspace{2em}
\hrule
\vspace{2em}
 \end{@twocolumnfalse}
  ]

@@

In the lower panel of Figure \ref{DailyAveragedDstEvents} it is shown that the trends seen in the analysis of 1989-1991 also hold for the entire span of measurements from GOES 6, which spans 1983-1991 - a small elevation on the day of the event relative to the day before and day after, with median averages after the day of the event being slightly lower than that before.  
To determine if the magnitude of the density medians in the top panel of \ref{DailyAveragedDstEvents} on days -1, 0, and 1 are statistically different from each other, we used a two-population bootstrap test.  A bootstrap median was created by sampling the values used to compute the median on each day with replacement.  The observed difference between the medians is shown in the table along with the fraction of observations in 1000 bootstrap differences with a different sign than that observed.  The results are shown in the fourth column of Table~\ref{BootstrapDifferenceTable}. The fourth column shows the results of a 2-population t-test and the fifth column indicates ???.

If a statistically significant difference in the medians between two days existed, the fraction is expected to be much smaller ($<$ 5\%).  In terms of a hypothesis test, we cannot reject the null hypothesis that the medians are the same with a confidence higher than 100 minus the \% value shown in the table; the highest confidence level of 87.6\% is in the difference between day 0 and day 1.

These tests indicate that there is not a statistically significant change in medians of $\rho_{eq}$ on the day before and day after a threshold crossing in $D_{st}$.  In addition, the observed differences are on the order of only 10\% of the $\rho_{eq}$, which is small compared to the variance in the observations.

%\input{../tables/DeltaBootstraps.txt}

\begin{table}
\small
\begin{tabular}{|c|cCc|}
 \hline 
Days & Observed diff & \%  & p-val \\ \hline
-1 0 & -0.64 & 28.00\%\geq 0 & 0.23 \\ 
-1 1 & 0.99 & 22.40\%\leq 0 & 0.86 \\ 
 1 0 & -1.63 & 12.40\%\geq 0 & 0.16 \\ 
\hline
\end{tabular}
\caption{Results of test on of means of $\rho_{eq}$ shown in the top panel of Figure~\ref{DailyAveragedDstEvents} between days of threshold crossing  near (day = 1 or -1) or on the day of a $D_{st}$ event (day = 0).}
\label{BootstrapDifferenceTable}
\end{table}
